---
title: "rcreds - Securely Incorporate Login Credentials Into R Scripts"
author: "Rick Saporta, Mike Reca"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


rcreds allows for a secure way to create R scripts that require login to other sytems, such as databases or websites. Users can securely write credentials and other sensitive information to an encrypted file on disk, and then later read those credentials back into R. 

The key concept has been adopted from 
      https://github.com/sdoyen/r_password_crypt/blob/master/crypt.R


## Overview

Many scripts require login to other systems.  The most common example is accessing data from a database. 
Including the login credentials in plain text in the script is a bad habbit that many of us implement when scheduling scripts or to avoid having to type in our password everytime we run the script. 

We will assume that, by nature of searching out such a package we can avoid here the discussion of the _why_ to implement a different approach (even when _"I am the only one who will be using or accessing this"_) and simply move on to the _how_ to use.

## Usage

Let's assume you have some function which requires the use of sensitive inputs. 

```{r, echo=FALSE, results='asis'}
some_login_function <- function(username, password, separate_param) {
     ## does something with username/password
     ## ... 

      message(sprintf("some_login_function() received username = '%s' and password = '%s' and separate_param = '%s'\n   (obviously wouldn't normally output like this)", username, password, separate_param))

     return(TRUE)
}
```
 
The `rcreds` approach would be to write the credentials to an encrypted file on disk and then read and decrypt them when needed. 

## Saving Credentials


One time, write to disk
```{r, echo=FALSE, results='asis'}

library(rcreds)

creds_info <- "for_app123_login" ## some description that will be part of the filename
key_object <- create_key(bytes=32, depth=8, verbose=TRUE)

## Save Credentials
write_credentials_to_file(username="cosmo", password="too many secrets", key=key_object, info.file_name = creds_info, verbose=TRUE)

## Save key file to a different location. 
save_key(key=key_object, zArchive_existing=FALSE, overwrite_existing=TRUE, verbose=TRUE)

```


Then in a script that needs to use the credentials, read from disk and use the list elements.
```{r, echo=FALSE, results='asis'}

key_file <- '~/.rcreds/key_files/.crypt_key.rds'
creds_file <- "~/.rcreds/credential_files/for_app123_login.credentials.creds"


creds <- read_credentials_from_file(file_full_path=creds_file, key=key_file)

## SHOWING CONTENTS FOR DEMO PURPOSES. NORMALLY DON'T DO THIS
print(creds)

## Use the credentials by refering to the elements by name
some_login_function(username = creds$username, password=creds$password, separate_param="plain example")

## altenatively can use do.call
do.call(some_login_function, c(creds, list(separate_param="do.call example")))
```



